---
title: "HMO changes over time"
author: "Jasmine Plows"
date: "4/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


```{r load_scripts, include = FALSE}
source('01-import.R')
source('02-tidy.R')
```

<b>1. Average concentrations of HMOs over time split by secretor status</b>
```{r}
source("HMO_changes_plotted_together.R")
plot_with_SUM
```

<b>2. Average concentrations of HMOs over time split by secretor status with SUM removed</b>
```{r}
source("HMO_changes_plotted_together.R")
plot_without_SUM
```

<b>3. Concentrations of HMOs over time showing each individual ALL DATA</b>
```{r, results='hide'}
source("HMO_changes_individuals_purrr.R")
all_individual_plots
```

<b>4. Concentrations of HMOs over time showing each individual COMPLETE DATA ONLY UP TO 12 MONTHS</b>
```{r, results='hide'}
source("HMO_changes_complete_obs.R")
all_individual_plots_complete_cases_12m
```

<b>5. Concentrations of HMOs over time showing each individual COMPLETE DATA ONLY UP TO 18 MONTHS</b>
```{r, results='hide'}
source("HMO_changes_complete_obs.R")
all_individual_plots_complete_cases_18m
```

<b>6.Concentrations of HMOs over time showing median and interquartile range ALL DATA</b>
```{r, results='hide'}
source("HMO_changes_median_IQR_plots.R")
all_plots
```

<b>7.Concentrations of HMOs over time showing median and interquartile range COMPLETE DATA ONLY UP TO 12 MONTHS</b>
```{r, results='hide'}
source("HMO_changes_median_complete_obs.R")
all_plots_median_complete_12m
```

<b>8.Concentrations of HMOs over time showing median and interquartile range COMPLETE DATA ONLY UP TO 18 MONTHS</b>
```{r, results='hide'}
source("HMO_changes_median_complete_obs.R")
all_plots_median_complete_18m
```